---
title: 'DATA607: Final Project Data Scraping Glassdoor'
author: "Eric Lehmphul"
date: "11/14/2021"
output: html_document
---

```{r}
library(tidyverse)
library(rvest)
library(xml2)
```

The data we need to scrape is on Glassdoor. The items we will acquire to use in our recommender system are: 

* education, base salary, avg rating, industry,
* company, job title, job description,

maybe rating as well


# Code

```{r}
url = "https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP1.htm"



page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)
```

## Functions to scrape desired data

### Job Description

```{r}
get_description <- function(job_link){
  job_page <- read_html(job_link)
  job_description <- job_page  %>% 
                      html_nodes(".ecgq1xb4") %>% html_text()
  
  return(job_description)
}
```

### Company Name

```{r}
get_company_name <- function(page){
  company <- page  %>% 
              html_nodes(".company span") %>% html_text()
  
  return(company)
}
```


### Location

```{r}
get_job_location <- function(page){
  location <- page  %>% 
              html_nodes(".location span") %>% html_text()
  
  return(location)
}
```


### Salary

```{r}
get_salary <- function(job_link){
  job_page <- read_html(job_link)
  job_salary<- job_page  %>% 
                        html_nodes(".e11nt52q2 .e1v3ed7e1") %>% html_text()
  
  return(job_salary)
}
```



### Average Company Rating

```{r}
get_rating <- function(job_link){
  job_page <- read_html(job_link)
  rating <- job_page  %>% 
                        html_nodes(".e11nt52q4") %>% html_text()
  
  return(rating)
}
```

### Industry of Company

```{r}
get_industry <- function(job_link){
  a <- readLines(job_link)
  value <- grep("industry",a)
  industry <- str_extract(a[value], "industryName\":\"(.*?)\"") %>%
    str_remove("industryName\":") %>%
    str_remove_all("\"")
    
  return(industry)
}
```


```{r}
# a <- readLines(job_link[1])
# 
# value <- grep("industry",a)
# 
# # a[value]
# 
# industry <- str_extract(a[value], "industryName\":\"(.*?)\"") %>%
#   str_remove("industryName\":") %>%
#   str_remove_all("\"")
```


### Sector of Company


Use similar approach to industry. look at source code of webpage and extract the words before the sector: sectorName

```{r}
get_sector <- function(job_link){
  a <- readLines(job_link)
  value <- grep("sector",a)
  sector <- str_extract(a[value], "sectorName\":\"(.*?)\"") %>%
    str_remove("sectorName\":") %>%
    str_remove_all("\"")
    
  return(sector)
}
```

### Job Title

```{r}
get_title <- function(page){
  title <- page  %>% 
                html_nodes(".job_title") %>% html_text()
  
  return(title)
}
```


### Company Reviews

```{r}
get_reviews <- function(job_link){
  job_page <- read_html(job_link)
  review <- job_page  %>% 
                        html_nodes(".ea78plz5") %>% html_text()
  
  return(review)
}
```


## Can be applied after scraping on the job description

### Education


```{r}
get_educaion_level <- function(dataset){
  # store descriptions
  description <- dataset$job_descripion
  
  # get true/false values for these key words 
  bachelors <- grepl("Bachelor|bachelor|BS|B.S.", ignore.case = F, b$job_description)
  masters <- grepl("Master|Graduate|master|MS|M.S.", ignore.case = F, b$job_description)
  phd <- grepl("Phd|doctorate|Postgraduate", ignore.case = T, b$job_description)
  
  # combine into a single list
  education <- cbind(bachelors, masters, phd)
  
  # convert TRUE to 1 and FALSE to 0
  education <- as.data.frame(ifelse(education == TRUE, 1, 0))
  
  # add education columns
  new.data <- cbind(dataset, education)
}
```





# Test

## Company Name

```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
page <- read_html(url)

descriptions <- get_company_name(page)
```

```{r}
descriptions[1:2]
```


## Job Description

```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
  page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

descriptions <- sapply(job_link, FUN = get_description)
```

```{r}
descriptions[1:2]
```

## Location

```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
page <- read_html(url)

descriptions <- get_job_location(page)
```

```{r}
descriptions[1:2]
```


## Job Salary

```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
  page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

descriptions <- sapply(job_link, FUN = get_salary)
```

```{r}
# Extract low and high salary range

salary = 0
for(j in 1:length(descriptions)){
salary[j] <- substr(descriptions[j], 1, 23)
}

salary[1:2]
```


### Company Rating

```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
  page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

descriptions <- sapply(job_link, FUN = get_rating)
```

```{r}
# Extract rating

rating = 0
for(j in 1:length(descriptions)){
rating[j] <- substr(descriptions[j], 1, 3)
}

rating[1:2]
```


### Company Industry

If company returns 'Industry' as value then the company did not post the industry it is in.
```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
  page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

descriptions <- sapply(job_link, FUN = get_industry)
```

```{r}
descriptions[1:5]
```

```{r}
job_link[5]
```

### Company Sector

If company returns 'Sector' as value then the company did not post the sector it is in.
```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
  page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

descriptions <- sapply(job_link, FUN = get_sector)
```

```{r}
descriptions[1:5]
```

```{r}
job_link[5]
```


### Job Title

```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
  page <- read_html(url)


descriptions <- get_title(page)
```

```{r}
descriptions[1:5]
```


### Company Reviews

```{r}
url = paste0("https://www.glassdoor.com/Job/new-york-data-scientist-jobs-SRCH_IL.0,8_IC1132348_KO9,23_IP", 1, ".htm")
  
  page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

descriptions <- sapply(job_link, FUN = get_reviews)
```

```{r}
descriptions[1:5]
```


## Test For Loop on 2 pages with functions created

```{r, warning=FALSE}
# initialize dataframe
scraped.data <- 0

for(page_result in 1:2){

url = paste0("https://www.glassdoor.com/Job/data-scientist-jobs-SRCH_KO0,14_IP", page_result, ".htm")
  
page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

# Normal function calls

company_name <- get_company_name(page)
job_location <- get_job_location(page)
job_title <- get_title(page)

# sapply function calls

job_description <- sapply(job_link, FUN = get_description)
company_industry <- sapply(job_link, FUN = get_industry)
company_sector <- sapply(job_link, FUN = get_sector)
company_rating <- sapply(job_link, FUN = get_rating)
job_salary <- sapply(job_link, FUN = get_salary)


dataset <- cbind(job_title, job_description, job_salary, job_location, company_name, company_industry, company_sector, company_rating)

scraped.data <- rbind(scraped.data, dataset)

print(paste("Page:", page_result))
}
```

Make data into a data frame and eliminate first row because it is a null row.

```{r}
final.data <- as.data.frame(scraped.data)
final.data <- final.data[-1,]
```

Add education qualifications in by scanning for key words in the job description. (bachelors, masters, phd| doctorate)

```{r}
final.data <- get_educaion_level(final.data)
```


```{r}
# clean up data 

# Rating

final.data$company_rating <- str_extract(final.data$company_rating, '\\d*\\.?\\d+')



# Salary

final.data$job_salary <- gsub(" \\..*", "", final.data$job_salary)
```


```{r}
# Change lists to proper data type

final.data <- as.data.frame(final.data)

final.data$job_title <- as.character(final.data$job_title)
final.data$job_description <- as.character(final.data$job_description)
final.data$job_salary <- as.character(final.data$job_salary)
final.data$job_location <- as.character(final.data$job_location)
final.data$company_name <- as.character(final.data$company_name)
final.data$company_industry <- as.character(final.data$company_industry)
final.data$company_sector <- as.character(final.data$company_sector)
final.data$company_rating <- as.character(final.data$company_rating)
```

```{r}
final.data$company_rating <- as.numeric(final.data$company_rating)
```


```{r}
final.data$company_industry[final.data$company_industry == "Industry"] <- "NA"
final.data$company_sector[final.data$company_sector == "Sector"] <- "NA"
```

Other data cleaning steps to be taken will be to separate the city and state, separate the upper and lower salary, and clean up company sector.


# Run Scraper to acquire full dataset

```{r, warning=FALSE}
# initialize dataframe
scraped.data <- 0

for(page_result in 1:15){

url = paste0("https://www.glassdoor.com/Job/data-scientist-jobs-SRCH_KO0,14_IP", page_result, ".htm")
  
page <- read_html(url)

job_link <- page  %>% 
  html_nodes(".jobLink") %>% html_attr("href") %>%
  paste("https://www.glassdoor.com", ., sep = "")

job_link <- unique(job_link)

# Normal function calls

company_name <- get_company_name(page)
job_location <- get_job_location(page)
job_title <- get_title(page)

# sapply function calls

job_description <- sapply(job_link, FUN = get_description)
company_industry <- sapply(job_link, FUN = get_industry)
company_sector <- sapply(job_link, FUN = get_sector)
company_rating <- sapply(job_link, FUN = get_rating)
job_salary <- sapply(job_link, FUN = get_salary)


dataset <- cbind(job_title, job_description, job_salary, job_location, company_name, company_industry, company_sector, company_rating)

scraped.data <- rbind(scraped.data, dataset)

print(paste("Page:", page_result))
}
```

Make data into a data frame and eliminate first row because it is a null row.

```{r}
final.data <- as.data.frame(scraped.data)
final.data <- final.data[-1,]
```

Add education qualifications in by scanning for key words in the job description. (bachelors, masters, phd| doctorate)

```{r}
final.data <- get_educaion_level(final.data)
```


```{r}
# clean up data 

# Rating

final.data$company_rating <- str_extract(final.data$company_rating, '\\d*\\.?\\d+')



# Salary

final.data$job_salary <- gsub(" \\..*", "", final.data$job_salary)
```


```{r}
# Change lists to proper data type

final.data <- as.data.frame(final.data)

final.data$job_title <- as.character(final.data$job_title)
final.data$job_description <- as.character(final.data$job_description)
final.data$job_salary <- as.character(final.data$job_salary)
final.data$job_location <- as.character(final.data$job_location)
final.data$company_name <- as.character(final.data$company_name)
final.data$company_industry <- as.character(final.data$company_industry)
final.data$company_sector <- as.character(final.data$company_sector)
final.data$company_rating <- as.character(final.data$company_rating)
```

```{r}
final.data$company_rating <- as.numeric(final.data$company_rating)
```


```{r}
final.data$company_industry[final.data$company_industry == "Industry"] <- "NA"
final.data$company_sector[final.data$company_sector == "Sector"] <- "NA"
```


# Export to .csv to share with team

```{r}
write.csv(final.data, "glassdoor_datascience.csv")
```
