---
title: 'Final Project: Job Recommender'
author: by Team Shiny (William Aiken, Bikram Barua, Catherine Cho, Eric Lehmphul, Nnaemeka Newman
  Okereafor)
date: "11/27/2021"
output: rmdformats::readthedown
---


### Introduction: Kaggle conducted a survey to gather information on the state of data science and machine learning around the world. Glassdoor data science job dataset .... Online resume dataset .... 

#### Purpose: In this project, we present a recommender system designed for job seeking and recruiting purpose. The proposed recommender system aims at leveraging the jobs and companies that are important for a target candidate and vice-versa. To meet this objective user profile such as job descriptions and candidate resumes are examined along with other user inputs. The recommendation approach is modeled on content-based analysis .... The dataset consisted of survey dataset of 23,859 respondents and job postings from [Glassdoor](http://www.glassdoor.com) and resumes from [Post Resumes Free](https://www.postjobfree.com/resume/adktqz/senior-data-scientist-brooklyn-ny) 

##### Research Question: How reliable is the job description in the listings that a candidate who had applied to the position in predicting the strength of their preference for the skills in the resume ? How reliable are company ratings in predicting candidates preference towards the job?


##### Approach: We have applied the following process to organize our workflow:

1) Business Understanding - TBDs
2) Data Understanding - ...
3) Data Preparation -  ...
4) Modeling - ...
5) Evaluation -  ...
6) Integration and Deployment - ...



############################################################################
# Instructions
This project is a proof-of-concept(POC) with certain assumptions on the data. For this implementation, following are the guidelines ...




# Load the data

```{r warning=FALSE, message=FALSE}


```


## Filter the datasets


```{r warning=FALSE, message=FALSE}

```




## Join the dataframes


```{r warning=FALSE, message=FALSE}

```


## Summaruze the data



# Data Modelling 
In our system, each entity has a profile. We use the content- and interaction-based relations to connect entities.

Profile match: Bidirectional relation based on the hypothesis that if a candidate’s resume matches a job description, they are probably interested in each other.

.....

## Calculate ratings based on ... algorithm


```{r warning=FALSE, message=FALSE}

```



## Training and test data


```{r warning=FALSE, message=FALSE}

```



# Build Recommender


## Connect the entities

## Find similarities 

## Examine the relationships



## Test the Recommender




# Create the Shiny UI


## Authenticate the user

## Gather Inputs


## Present visualizations




# Conclusion

Evaluation: Our recommender predicted ...

Conclusion: ...

# Shiny Apps Links

# Q & A



